---
title: "STA238 Week 3 - Statistical Models"
author: "Samantha-Jo Caetano"
date: "January 25, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Linear Regression

Let's perform a linear regression analysis using the coffee ratings data that was introduced in week 1. 

### Load in the coffee ratings data here:
```{r coffee-load, include=FALSE}
library(tidyverse)
coffee_ratings <- read_csv("coffee_ratings.csv")
```


Let's start off by plotting some variables against each other to see what might make for a good linear relationship (we want something appropriate both practically and statistically).


```{r scatterplots, echo=FALSE}


```


So our model is as follows: $$Y_i = \alpha + \beta x_i + U_i$$ or (some notation also use $$Y_i = \beta_0 + \beta_1 x_i + \epsilon_i$$). Later on we will learn about how to estimate these parameters, but for now let's let R do the work for us.

```{r slope-intercept-est, echo=FALSE}



```


Now let's re-plot the original the original scatterplot but with the affiliated line.


```{r scatterplots-with-line, echo=FALSE}




```



## Extra

```{r Quadratic Estimation, echo=FALSE}




```




# Unbiasedness


Let's revisit last week's example. Assume $$X_1, ..., X_n ~ Exp(\beta = 0.5)$$, where $$\beta$$ is the rate parameter. Thus, $$E(X) = 1/\beta = 2$$ and $$Var(X) = 1/\beta^2 = 4 $$. Now let's use a simulation to "show" that the sample variance is unbiased.

```{r}
set.seed(20210126)

# Simulate 500 random samples of size 30 from a Exp(1/2)
N <- 500
n <- 30
beta <- 0.5

# Simulate the samples and calculate the estimators for each sample
samples <- vector(mode = "list",length = N)

VARs_from_Samples <- numeric(N) # Create an empty vector - one placeholder for each of the N samples.
for (i in 1:N) {
  samples[[i]] <- rexp(n, beta) 
  VARs_from_Samples[i] <- var(samples[[i]])
}

tibble(VARs_from_Samples) %>%
  ggplot(aes(x = VARs_from_Samples)) +
  theme_classic() +
  geom_histogram(colour = "black",fill = "transparent",bins = 10) +
  coord_cartesian(ylim = c(0,250)) +
  geom_vline(xintercept = 4,colour = "red",linetype = "dotdash")

mean(VARs_from_Samples)


```











